---
title: "SMPA - Grouping Assignment"
author: "YourNameHere"
date: "`r format(Sys.time(), '%B %d, %Y')`"
format:
  html:
    self-contained: true
---

```{r}

#load libraries
library(tidyverse)
library(lubridate)

#load data
payments <- readRDS("data/trump_campaign_payments.rds")

```

## Introduction

**What data will you be working with?**

A dataset of disbursements (payments and purchases) by the Donald J. Trump for President campaign committee during the 2020 election cycle. The included payments are from 2017 and 2018.

This data and examples were developed as part of the NYT Open training, which we've adapted from here. Note that the dataset has been alterted for instructional purposes, so doesn't represent the original FEC data in its entirety.

In the environment tab in the upper right, click on the `payments` dataset to open the RStudio viewer to examine the data. You can also run the chunk below as well to bring up a slice of the data.

```{r}

head(payments)

##test

```

## Before you begin

Take a moment to look at the various categories included for each disbursement.

-   Notice that there are several date-related fields:

    -   FEC election cycle

    -   report year

    -   election type

    -   disbursement date

-   There are a few fields about the recipient. Along with city and state, you have:

    -   the recipient name

    -   the payee's first and last name (an individual who was paid for the "recipient") and

    -   the entity type (individual v. organization)

## Initial Narrative Questions

Without doing any analysis yet, what are three questions you might try to answer using this dataset? Write your narrative answers in the spaces below.

1.  What state did the campaign committee spend the most on? This dataset includes both states and disbursement amounts 

2.  What year had the highest disbursement rates? This dataset includes both year and a total number of disbursements, so it'd be interesting to see what year was most focused on.

3.  what category did the campaign spend the most on in Virginia? It'd be interesting to see what the committee prioritized close to home!

## Coding Assignment Questions

**Please write your R tidyverse code to help answer the questions below in the provided R chunks. You can make more than one chunk if you prefer.**

**Also, describe what you're doing either in text out "in the open" above the chunk, or in comments (prefaced by #) in the R chunk itself.**

**You'll need to use grouping, filtering, or a combination of both to find the answers. Remember that R is case-sensitive when you refer to any text values within the columns**

<br>

How much was spent in the 2017 vs. 2018 reporting year?

```{r}

payments2017 <-
  filter(payments, disbursement_year == 2017) %>%
  group_by(committee_name) %>% 
  summarise(total_disbursement = sum(disbursement_amount))
  ##17,082,338

payments2018 <-
  filter(payments, disbursement_year == 2018) %>%
  group_by(committee_name) %>% 
  summarise(total_disbursement = sum(disbursement_amount))
  ##37,212,716
  






```

Tally the total overall disbursement amounts for each month (distinguishing between say, Oct. 2018 v. Oct. 2019). In which month was spending the highest?

```{r}

paymentsmonths <-
  payments %>% 
  group_by(committee_name, disbursement_year, disbursement_month) %>% 
  summarise(monthly_disbursement = sum(disbursement_amount)) 


##October of 2018

```

Which recipient (recipient_name) was paid the most in the 2020 election cycle?

```{r}

toprecipient2020 <-
  filter(payments, election_cycle == 2020) %>% 
  group_by(recipient_name) %>% 
  summarise(totaldisbursement = sum(disbursement_amount)) 


##American Made Media Consultants LLC



```

Look at what this recipient was paid for. What service or good (see the "disbursement_description") were they paid the most for?

```{r}

mediaconsultants <-
  filter(payments, recipient_name == "AMERICAN MADE MEDIA CONSULTANTS, LLC", election_cycle == 2020) %>% 
  group_by(recipient_name, disbursement_description) %>% 
  summarise(totaldisbursement = sum(disbursement_amount)) 
##Placed media



```

What's the most common category overall (see "disbursement_description") the campaign spent money on?

```{r}
commoncategory <-
  payments %>% 
  group_by(committee_name, disbursement_description) %>% 
  summarise(totaldisbursement = sum(disbursement_amount))

##DIGITAL CONSULTING/ONLINE ADVERTISING



```

Who were the 5 biggest recipients of campaign spending on "LEGAL CONSULTING"..?

```{r}


legalconsulting <- 
  filter(payments, disbursement_description == "LEGAL CONSULTING") %>% 
  group_by(recipient_name, disbursement_description) %>% 
  summarise(totaldisbursement = sum(disbursement_amount)) %>% 
  arrange(desc(totaldisbursement)) %>% 
  head(5) %>% 
  view()

##See view

```


Who are the top 5 individuals (vs. organization) were paid by the campaign committee?

```{r}


top5individual <-
  payments %>%  
  filter(payee_last_name != "NA") %>% 
  group_by(payee_last_name, payee_first_name) %>% 
  summarise(totaldisbursement = sum(disbursement_amount)) %>%  
  arrange(desc(totaldisbursement)) %>% 
  head(5) %>% 
  view() 


##see view




```


For the indvidual who was paid the most, which payment category (disbursement_description) was the top one for him or her? And how much money were they paid in that category?

```{r}

glassneramounts <-
  payments %>% 
  filter(payee_last_name == "GLASSNER") %>% 
  group_by(payee_last_name, disbursement_description) %>%
  summarise(totaldisbursement = sum
            (disbursement_amount))

##Payroll, $499,000



```
